## This script is to explain how intermediate files were prepared.
## Some intermediate files were either large or would take a long time to generate


Figure 1. 

List of intermediate files
1.1) four_organ_gene_exp_filtered.rds: gene counts with extremely low coverage is removed.
1.2) four_organs_gene_reformatted_v2.rds: gene counts reformated for downstream analysis
1.3) four_organs_lncRNA_exp_raw_normalized.rds: 
1.4) four_organs_lncRNA_exp_raw_normalized_v2.rds
1.5) four_organ_TE_subfamily_exp_filtered.rds
1.6) four_organs_TE_subfamily_reformatted_v2.rds
1.7) heatmap_four_organ_normalized.rds
1.8) Figure_1c_normalized_all_TEinstance.rds



### Figure 1.

#### Libraries and global variables
```{r}
## Libraries
suppressMessages(library(tidyverse))
suppressMessages(library( "DESeq2" ))
suppressMessages(library(ggplot2))
suppressMessages(library(ggfortify))
suppressMessages(library(readxl))
suppressMessages(library(dplyr))

## global variable
dir_github_script <- "G:/My Drive/github_script/" ## replace with your local directory

```


#### Intermediate file 1.1) four_organ_gene_exp_filtered.rds
#### Intermediate file 1.2) four_organs_gene_reformatted_v2.rds
```{r}
# dir_github_script <- "G:/My Drive/github_script/"
dir_github_intermediate_files <- paste0(dir_github_script, "intermediate_files_prep/Figure1/collection_intermediate_files/")

dir_TE_exp <- paste0(dir_github_script, "intermediate_files_prep/Figure1/")
dir_gene_exp <- paste0(dir_github_script, "intermediate_files_prep/Figure1/")
gene_chr_pos <- read.csv(file = paste0(dir_TE_exp, "mm10_gene_loc_uniq.csv"), header = TRUE)

#### Four organ gene exp PCA
four_organ_gene_exp <- read.csv(paste0(dir_gene_exp, "Four_organs_gene_tb_raw.csv"))
#### To include HXX14 that was missing. 
heart_HXX14f <- read_excel(paste0(dir_gene_exp, "new_HXX14_counts.xlsx"))

#### Remove genes with extremely low coverage: read sums < 50 
four_organ_gene_exp_filtered <- four_organ_gene_exp[rowSums(four_organ_gene_exp[, 3:73 ])> 50,]

four_organ_gene_exp_filtered <- left_join(four_organ_gene_exp_filtered, gene_chr_pos)  %>% na.omit()

Sample.ID <- four_organ_gene_exp_filtered[,3:73] %>% names()
Organ <- c(rep("Liver", 21), rep("Brain", 18), rep("Heart", 16), rep("Lung", 16))
Group <- c(rep("XX.F", 4), rep("XY.M", 4), rep("XY.F", 4), rep("XO.F", 4), rep("XXpaf.F", 5),
           rep("XX.F", 4), rep("XY.M", 4), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 3),
           rep("XX.F", 2), rep("XY.M", 3), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 4),
           rep("XX.F", 3), rep("XY.M", 3), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 3))
gene.TE <- four_organ_gene_exp_filtered$Gene

four_organs_colors <- c("#FF7F0E", "#2CA02C", "#1F77B4", "#e377c2")

#### raw counts
four_organ_normalized_raw <- four_organ_gene_exp_filtered[, 3:73]
#### update to be norm counts, normalized by library depth of each sample.
four_organ_sum_depths <- four_organ_normalized_raw %>% lapply(., sum) %>% unlist()
four_organ_sum_depths_average <- four_organ_sum_depths  %>% mean()
for(i in 1:70){
  four_organ_normalized_raw[, i] <- (four_organ_normalized_raw[, i] * four_organ_sum_depths_average/four_organ_sum_depths[i]) %>% as.integer()
}

## To include chromosome information for gene
Gene_chr <- four_organ_gene_exp_filtered$chr
four_organ_normalized_raw_labeled <- cbind(four_organ_normalized_raw, Gene_chr)
#### X-linked
four_organ_normalized_raw_X <- four_organ_normalized_raw_labeled %>% filter(Gene_chr == "chrX") %>% dplyr::select(!"Gene_chr")
#### Autosomal
four_organ_normalized_raw_auto <- four_organ_normalized_raw_labeled %>% filter(Gene_chr != "chrX",
                                                                               Gene_chr != "chrY",
                                                                               Gene_chr != "chr1_GL456221_random",
                                                                               Gene_chr != "chr4_JH584293_random",
                                                                               Gene_chr != "chr4_JH584294_random",
                                                                               Gene_chr != "chrX_GL456233_random",
                                                              !is.na(Gene_chr)) %>% dplyr::select(!"Gene_chr")

### reformat the data for downstream analysis
four_organs_gene_reformatted_v2 <- cbind((t(four_organ_normalized_raw) ) %>% as_tibble(), 
                                         Sample.ID,
                                         Group,
                                         Organ) %>% data.frame()
four_organs_gene_reformatted_v2_auto <- cbind((t(four_organ_normalized_raw_auto) ) %>% as_tibble(), 
                                              Sample.ID,
                                              Group,
                                              Organ) %>% data.frame()
four_organs_gene_reformatted_v2_X <- cbind((t(four_organ_normalized_raw_X) ) %>% as_tibble(), 
                                           Sample.ID,
                                           Group,
                                           Organ) %>% data.frame()


rownames(four_organs_gene_reformatted_v2) <- four_organs_gene_reformatted_v2$Sample.ID


saveRDS(four_organ_gene_exp_filtered, file = paste0(dir_github_intermediate_files, "four_organ_gene_exp_filtered.rds"))
saveRDS(four_organs_gene_reformatted_v2, file = paste0(dir_github_intermediate_files, "four_organs_gene_reformatted_v2.rds"))

```



#### Intermediate file 1.3) four_organs_lncRNA_exp_raw_normalized.rds
#### Intermediate file 1.4) four_organs_lncRNA_exp_raw_normalized_v2.rds

```{r}
# dir_github_script <- "G:/My Drive/github_script/"
dir_lncRNA_count <- paste0(dir_github_script, "intermediate_files_prep/Figure1/")
dir_github_intermediate_files <- paste0(dir_github_script, "intermediate_files_prep/Figure1/collection_intermediate_files/")

#### read-in formated table of de novo lncRNA counts 
Four_organs_lncRNA_candidates_v2 <- readRDS(file = paste0(dir_lncRNA_count, 
                                                          "Four_organs_lncRNA_candidates_v2.rds"))

#### Four organs - lncRNA expression PCA
four_organs_lncRNA_exp <- Four_organs_lncRNA_candidates_v2
four_organs_lncRNA_exp_filtered <- four_organs_lncRNA_exp[rowSums(four_organs_lncRNA_exp[, 2:72])>50, ]

#### reformating for downstream analysis
all_Sample.ID <- c(four_organs_lncRNA_exp_filtered[,2:72] %>% names())
Organ <- c(rep("Liver", 21), rep("Brain", 18), rep("Heart", 16), rep("Lung", 16))
Group <- c(rep("XX.F", 4), rep("XY.M", 4), rep("XY.F", 4), rep("XO.F", 4), rep("XXpaf.F", 5),
           rep("XX.F", 4), rep("XY.M", 4), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 3),
           rep("XX.F", 3), rep("XY.M", 3), rep("XY.F", 4), rep("XO.F", 2), rep("XXpaf.F", 4),
           rep("XX.F", 3), rep("XY.M", 3), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 3))

four_organs_lncRNA_index <- four_organs_lncRNA_exp_filtered$index

four_organs_lncRNA_setup <- tibble(Sample.ID = all_Sample.ID,
                                           Group = Group,
                                           Organ = Organ)

### Normalization
four_organs_lncRNA_exp_raw <- four_organs_lncRNA_exp_filtered[, 2:72] %>% data.frame()
rownames(four_organs_lncRNA_exp_raw) <- four_organs_lncRNA_index


four_organs_lncRNA_exp_sum_depths <- four_organs_lncRNA_exp_raw %>% lapply(., sum) %>% unlist()
four_organs_lncRNA_exp_sum_depths_average <- four_organs_lncRNA_exp_sum_depths  %>% mean()

four_organs_lncRNA_exp_raw_normalized <- four_organs_lncRNA_exp_raw

for(i in 1:71){
  four_organs_lncRNA_exp_raw_normalized[, i] <- (four_organs_lncRNA_exp_raw[, i] * four_organs_lncRNA_exp_sum_depths_average/four_organs_lncRNA_exp_sum_depths[i]) %>% as.integer()
}

four_organs_lncRNA_exp_raw_normalized_v2 <- cbind((t(four_organs_lncRNA_exp_raw_normalized))%>%
                                                    as_tibble(), 
                                         all_Sample.ID,
                                         Group,
                                         Organ) %>% data.frame()

## save intermediate files
saveRDS(four_organs_lncRNA_exp_raw_normalized, file = paste0(dir_github_intermediate_files,
                                                             "four_organs_lncRNA_exp_raw_normalized.rds"))
saveRDS(four_organs_lncRNA_exp_raw_normalized_v2, file = paste0(dir_github_intermediate_files, 
                                                                "four_organs_lncRNA_exp_raw_normalized_v2.rds"))


```


#### Intermediate file 1.5) four_organ_TE_subfamily_exp_filtered.rds
#### Intermediate file 1.6) four_organs_TE_subfamily_reformatted_v2.rds
```{r}
# dir_github_script <- "G:/My Drive/github_script/"
dir_TE_exp <- paste0(dir_github_script, "intermediate_files_prep/Figure1/")
dir_github_intermediate_files <- paste0(dir_github_script, "intermediate_files_prep/Figure1/collection_intermediate_files/")

four_organ_TE_subfamily_exp <- read.csv(paste0(dir_TE_exp, "four_organ_TEfamily_only_exp.csv"))

four_organ_TE_subfamily_exp_filtered <- four_organ_TE_subfamily_exp[rowSums(four_organ_TE_subfamily_exp[, 2:69 ])> 50,]


Sample.ID <- four_organ_TE_subfamily_exp_filtered[,2:69] %>% names()
Organ <- c(rep("Brain", 17), rep("Heart", 15), rep("Lung", 16), rep("Liver", 20))
Group <- c(rep("XX.F", 3), rep("XY.M", 4), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 3),
           rep("XX.F", 2), rep("XY.M", 2), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 4),
           rep("XX.F", 3), rep("XY.M", 3), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 3),
           rep("XX.F", 4), rep("XY.M", 3), rep("XY.F", 4), rep("XO.F", 4), rep("XXpaf.F", 5))
gene.TE <- four_organ_TE_subfamily_exp_filtered$gene.TE

rownames(four_organ_TE_subfamily_exp_filtered) <- four_organ_TE_subfamily_exp_filtered$gene.TE
four_organ_normalized_raw <- four_organ_TE_subfamily_exp_filtered[,2:69]


four_organ_sum_depths <- four_organ_normalized_raw %>% lapply(., sum) %>% unlist()
four_organ_sum_depths_average <- four_organ_sum_depths  %>% mean()

for(i in 1:68){
  four_organ_normalized_raw[, i] <- (four_organ_normalized_raw[, i] * four_organ_sum_depths_average/four_organ_sum_depths[i]) %>%
    as.integer()
}

four_organs_TE_subfamily_reformatted <- cbind((t(four_organ_TE_subfamily_exp_filtered[,2:69] ) %>% as_tibble()), 
                                              Sample.ID,
                                              Group,
                                              Organ) %>% data.frame()

four_organs_TE_subfamily_reformatted_v2 <- cbind((t(four_organ_normalized_raw) ) %>% as_tibble(), 
                                                 Sample.ID,
                                                 Group,
                                                 Organ) %>% data.frame()

rownames(four_organs_TE_subfamily_reformatted) <- four_organs_TE_subfamily_reformatted$Sample.ID

rownames(four_organs_TE_subfamily_reformatted_v2) <- four_organs_TE_subfamily_reformatted_v2$Sample.ID
# colnames(four_organs_TE_subfamily_reformatted_v2) <- c(gene.TE, "Sample.ID", "Group", "Organ")


saveRDS(four_organ_TE_subfamily_exp_filtered, file = paste0(dir_github_intermediate_files, 
                                    "four_organ_TE_subfamily_exp_filtered.rds"))
saveRDS(four_organs_TE_subfamily_reformatted_v2, file = paste0(dir_github_intermediate_files, 
                                                               "four_organs_TE_subfamily_reformatted_v2.rds"))


```



#### Intermediate file 1.7) heatmap_four_organ_normalized.rds
```{r}
#### Intermediate file heatmap_four_organ_normalized.rds was prepared on cluster due to high computing demand
heatmap_four_organ_normalized <- readRDS(file = paste0(four_organ_TE_subfamily_exp_filtered,
                                                            "heatmap_four_organ_normalized.rds"))

```



#### Intermediate file 1.8) Figure_1c_normalized_all_TEinstance.rds
```{r}
#### Intermediate file Figure_1c_normalized_all_TEinstance.rds was prepared on cluster due to high computing demand
#### Figure 1C
Figure_1c_normalized_all_TEinstance <- readRDS(file = paste0(four_organ_TE_subfamily_exp_filtered,
                                                        "Figure_1c_normalized_all_TEinstance.rds"))


```






