## This script is to explain how intermediate files were prepared.
## Some intermediate files had a high demand for computing resources and were 
### separately prepared on cluster instead.
#### For the naming of variables, "_Both" is used as the abbreviation of "_XXfvsXYm"
#### For the naming of variables, "_Pheno" is used as the abbreviation of "_XYfvsXYm"


Figure 3. 

List of intermediate files
3.1) sDETE_to_sDETE_distance_Both.rds
3.2) sDETE_to_sDETE_distance_Pheno.rds
3.3) nonsDETE_to_sDETE_distance_Both.rds
3.4) nonsDETE_to_sDETE_distance_Pheno.rds
3.5) DEG_to_DEG_distance_Both.rds
3.6) DEG_to_DEG_distance_Pheno.rds
3.7) nonDEG_to_DEG_distance_Both.rds
3.8) nonDEG_to_DEG_distance_Pheno.rds
3.9) slncRNA_to_slncRNA_distance_Both.rds
3.10) slncRNA_to_slncRNA_distance_Pheno.rds
3.11) nonslncRNA_to_slncRNA_distance_Both_reformated.rds
3.12) nonslncRNA_to_slncRNA_distance_Pheno_reformated.rds
3.13) all_DETE_DEG_distance_Both.rds
3.14) all_DETE_DEG_distance_Pheno.rds
3.15) random_draft_DETE_to_gene_distance_Both.rds
3.16) random_draft_DETE_to_gene_distance_Pheno.rds
3.17) slncRNA_to_sDETE_distance_Both.rds
3.18) slncRNA_to_sDETE_distance_Pheno.rds
3.19) random_draft_DETE_to_lncRNA_distance_Both.rds
3.20) random_draft_DETE_to_lncRNA_distance_Pheno.rds


### Figure 3.
#### Libraries and global variables
```{r}
## Library
suppressMessages(library(tidyverse))
suppressMessages(library(readxl))
suppressMessages(library(dplyr))


## global variable
dir_github_script <- "G:/My Drive/github_script/" ## replace with your local directory
dir_TE_exp <- "G:/My Drive/wrapup/Brain_paper/Four_organ/result_excels/" 
dir_github_intermediate_files_fig3 <- paste0(dir_github_script, 
                                             "intermediate_files_prep/Figure3/collection_intermediate_files/")


## global function
check_region_region_distance <- function(tb1, tb2, compared_to_itself = FALSE){
  collector_tb1 <- NULL
  collector_tb2 <- NULL
  list_of_chr <- paste0("chr", c(1:19, "X", "Y"))

  for(i in list_of_chr){
      tmp_tb1 <- tb1 %>% filter(chr == i) %>% mutate(dist_check = NULL)
      tmp_tb2 <- tb2 %>% filter(chr == i) %>% mutate(dist_check = NULL)
      if(dim(tmp_tb1)[[1]] > 0 && dim(tmp_tb2)[[1]] > 0 && dim(tmp_tb1)[[1]] > 1){
          for(j in 1:dim(tmp_tb1)[[1]]){
            dist_start_start <- tmp_tb1[j, "start"][[1]] - tmp_tb2$start
            dist_start_end <- tmp_tb1[j, "start"][[1]] - tmp_tb2$end
            dist_end_start <- tmp_tb1[j, "end"][[1]] - tmp_tb2$start
            dist_end_end <- tmp_tb1[j, "end"][[1]] - tmp_tb2$end
            
            if(compared_to_itself){
              ## allowing comparing the list to itself by removing the identical region in counting distance
              index_to_not_identical_v1 <- which(dist_start_start != 0)
              index_to_not_identical_v2 <- which(dist_end_end != 0)
              index_to_not_identical <- intersect(index_to_not_identical_v1, index_to_not_identical_v2)
              dist_start_start <- dist_start_start[index_to_not_identical]
              dist_start_end <- dist_start_end[index_to_not_identical]
              dist_end_start <- dist_end_start[index_to_not_identical]
              dist_end_end <- dist_end_end[index_to_not_identical]

            }
            
            
            dist_start_start <- dist_start_start %>% as_tibble() %>% mutate(value = as.double(value))
            dist_start_end <- dist_start_end %>% as_tibble() %>% mutate(value = as.double(value))
            dist_end_start <- dist_end_start %>% as_tibble() %>% mutate(value = as.double(value))
            dist_end_end <- dist_end_end %>% as_tibble() %>% mutate(value = as.double(value))

            min_check1 <- min(dist_start_start * dist_start_end)
            min_check2 <- min(dist_end_start * dist_end_end)
            min_check <- min(min_check1, min_check2)
            if(min_check<=0){
              tmp_tb1[j, "dist_check"] <- 0
            } else if (min_check>0){
              tmp_tb1[j, "dist_check"] <- min(abs(dist_start_start), abs(dist_start_end),
                                              abs(dist_end_start), abs(dist_end_end))

              
            }
            
            # dist1 <- tmp_tb1[j, "pos"][[1]] - tmp_tb2$start
            # dist2 <- tmp_tb1[j, "pos"][[1]] - tmp_tb2$end
            # dist1 <- dist1 %>% as_tibble() %>% mutate(value = as.double(value))
            # dist2 <- dist2 %>% as_tibble() %>% mutate(value = as.double(value))
            # min_check <- min(dist1 * dist2)
            # if(min_check<=0){
            #   tmp_tb1[j, "dist_check"] <- 0
            # } else if (min_check>0){
            #   tmp_tb1[j, "dist_check"] <- min(abs(dist1), abs(dist2))
            # }
          }
        
        for(k in 1:dim(tmp_tb2)[[1]]){
            dist_start_start <- tmp_tb2[k, "start"][[1]] - tmp_tb1$start
            dist_start_end <- tmp_tb2[k, "start"][[1]] - tmp_tb1$end
            dist_end_start <- tmp_tb2[k, "end"][[1]] - tmp_tb1$start
            dist_end_end <- tmp_tb2[k, "end"][[1]] - tmp_tb1$end

            
            if(compared_to_itself){
              ## allowing comparing the list to itself by removing the identical region in counting distance
              index_to_not_identical_v1 <- which(dist_start_start != 0)
              index_to_not_identical_v2 <- which(dist_end_end != 0)
              index_to_not_identical <- intersect(index_to_not_identical_v1, index_to_not_identical_v2)
              dist_start_start <- dist_start_start[index_to_not_identical]
              dist_start_end <- dist_start_end[index_to_not_identical]
              dist_end_start <- dist_end_start[index_to_not_identical]
              dist_end_end <- dist_end_end[index_to_not_identical]

            }
            
            dist_start_start <- dist_start_start %>% as_tibble() %>% mutate(value = as.double(value))
            dist_start_end <- dist_start_end %>% as_tibble() %>% mutate(value = as.double(value))
            dist_end_start <- dist_end_start %>% as_tibble() %>% mutate(value = as.double(value))
            dist_end_end <- dist_end_end %>% as_tibble() %>% mutate(value = as.double(value))
            min_check1 <- min(dist_start_start * dist_start_end)
            min_check2 <- min(dist_end_start * dist_end_end)
            min_check <- min(min_check1, min_check2)
            if(min_check<=0){
              tmp_tb2[k, "dist_check"] <- 0
            } else if (min_check>0){
              tmp_tb2[k, "dist_check"] <- min(abs(dist_start_start), abs(dist_start_end),
                                              abs(dist_end_start), abs(dist_end_end))
              

            }

            # dist1 <- tmp_tb2[k, "start"][[1]] - tmp_tb1$pos
            # dist2 <- tmp_tb2[k, "end"][[1]] - tmp_tb1$pos
            # dist1 <- dist1 %>% as_tibble() %>% mutate(value = as.double(value))
            # dist2 <- dist2 %>% as_tibble() %>% mutate(value = as.double(value))
            # min_check <- min(dist1 * dist2)
            # if(min_check<=0){
            #   tmp_tb2[k, "dist_check"] <- 0
            # } else if (min_check>0){
            #   tmp_tb2[k, "dist_check"] <- min(abs(dist1), abs(dist2))
            # }
        }
          collector_tb1 <- rbind(collector_tb1, tmp_tb1)
          collector_tb2 <- rbind(collector_tb2, tmp_tb2)
      }
  }
  return(list(collector_tb1, collector_tb2))
}



```

### sDETE and sDETE
#### Intermediate file 3.1) sDETE_to_sDETE_distance_Both.rds
#### Intermediate file 3.2) sDETE_to_sDETE_distance_Pheno.rds
#### Intermediate file 3.3) nonsDETE_to_sDETE_distance_Both.rds
#### Intermediate file 3.4) nonsDETE_to_sDETE_distance_Pheno.rds

```{r}
# dir_github_script <- "G:/My Drive/github_script/" ## replace with your local directory
dir_TE_exp_fig3 <- paste0(dir_github_script, "intermediate_files_prep/Figure3/")
dir_to_Liver_TE_exp <- dir_TE_exp_fig3

### Check sDETE colocalizing with peer sDETE
Liver_Both_res3 <- readRDS(file = paste0(dir_to_Liver_TE_exp, "Liver_Both_res3.rds")) %>% 
  separate(chromosome.start.stop, c("chr", "start", "end"))%>%
  mutate(chr = factor(chr), start = as.numeric(start), end = as.numeric(end)) %>% 
  filter(!is.na(start))
Liver_Pheno_res3 <- readRDS(file = paste0(dir_to_Liver_TE_exp, "Liver_Pheno_res3.rds"))%>% 
  separate(chromosome.start.stop, c("chr", "start", "end"))%>%
  mutate(chr = factor(chr), start = as.numeric(start), end = as.numeric(end))%>% 
  filter(!is.na(start))

Liver_Both_sDETE <- Liver_Both_res3 %>% filter(padj < 0.05, abs(log2FoldChange) > 1)
Liver_Pheno_sDETE <- Liver_Pheno_res3 %>% filter(padj < 0.05, abs(log2FoldChange) > 1)
Liver_Both_nonsDETE <- left_join(Liver_Both_res3, Liver_Both_sDETE %>% mutate(sDETE = "Yes")) %>% 
                          filter(is.na(sDETE))
Liver_Pheno_nonsDETE <- left_join(Liver_Pheno_res3, Liver_Pheno_sDETE %>% mutate(sDETE = "Yes")) %>% 
                          filter(is.na(sDETE))

sDETE_to_sDETE_distance_Both <- check_region_region_distance(Liver_Both_sDETE, 
                                                      Liver_Both_sDETE, compared_to_itself = TRUE)
sDETE_to_sDETE_distance_Pheno <- check_region_region_distance(Liver_Pheno_sDETE, 
                                                         Liver_Pheno_sDETE, compared_to_itself = TRUE)
### nonsDETE_to_sDETE_distance_Both and nonsDETE_to_sDETE_distance_Pheno takes a long time to generate
# nonsDETE_to_sDETE_distance_Both <- check_region_region_distance(Liver_Both_nonsDETE,
#                                                       Liver_Both_sDETE, compared_to_itself = TRUE)
# nonsDETE_to_sDETE_distance_Pheno <- check_region_region_distance(Liver_Pheno_nonsDETE,
#                                                       Liver_Pheno_sDETE, compared_to_itself = TRUE)


saveRDS(sDETE_to_sDETE_distance_Both, file = paste0(dir_github_intermediate_files_fig3, "sDETE_to_sDETE_distance_Both.rds"))
saveRDS(sDETE_to_sDETE_distance_Pheno, file = paste0(dir_github_intermediate_files_fig3, "sDETE_to_sDETE_distance_Pheno.rds"))
saveRDS(nonsDETE_to_sDETE_distance_Both, file = paste0(dir_github_intermediate_files_fig3, "nonsDETE_to_sDETE_distance_Both.rds"))
saveRDS(nonsDETE_to_sDETE_distance_Pheno, file = paste0(dir_github_intermediate_files_fig3, "nonsDETE_to_sDETE_distance_Pheno.rds"))

```


### sDEG and sDEG
#### Intermediate file 3.5) DEG_to_DEG_distance_Both.rds
#### Intermediate file 3.6) DEG_to_DEG_distance_Pheno.rds
#### Intermediate file 3.7) nonDEG_to_DEG_distance_Both.rds
#### Intermediate file 3.8) nonDEG_to_DEG_distance_Pheno.rds

```{r}
# dir_github_script <- "G:/My Drive/github_script/" ## replace with your local directory
dir_TE_exp_fig3 <- paste0(dir_github_script, "intermediate_files_prep/Figure3/")
dir_to_Liver_TE_exp <- dir_TE_exp_fig3
dir_to_fourOrgan_TE_subfamily_exp <- dir_TE_exp_fig3

## Read in DEG files in pairwise comparisons 
Liver_Both_res3 <- readRDS(file = paste0(dir_to_Liver_TE_exp, "Liver_Both_res3.rds"))
Liver_Pheno_res3 <- readRDS(file = paste0(dir_to_Liver_TE_exp, "Liver_Pheno_res3.rds"))
pos_of_gene_dedup <- readRDS(file = paste0(dir_to_fourOrgan_TE_subfamily_exp, "pos_of_gene_dedup.rds"))


Liver_DEG_list <- read_excel(paste0(dir_TE_exp_fig3, "Liver_DEG_recompiled.xlsx")) %>% mutate(gene = gene_symbol,
                                                                                         DEG = "Yes") %>% 
                                                                                  filter(abs(log_FC) > 1, log_CPM > 1, 
                                                                                         deseq.adj.pvalue < 0.05)
Liver_DEG_Both <- Liver_DEG_list %>% filter(Comparison == "XX.F vs XY.M")
Liver_DEG_Pheno <- Liver_DEG_list %>% filter(Comparison == "XY.F vs XY.M")


Liver_DEG_list_Both_pos <- left_join(Liver_DEG_Both, pos_of_gene_dedup %>% dplyr::rename(gene_symbol = gene))
Liver_DEG_list_Pheno_pos <- left_join(Liver_DEG_Pheno, pos_of_gene_dedup %>% dplyr::rename(gene_symbol = gene))

Liver_nonDEG_list_Both_pos <- (left_join(pos_of_gene_dedup %>% dplyr::rename(gene_symbol = gene), Liver_DEG_Both) %>% 
                                          filter(is.na(Comparison )))[, 1:4]
Liver_nonDEG_list_Pheno_pos <- (left_join(pos_of_gene_dedup %>% dplyr::rename(gene_symbol = gene), Liver_DEG_Pheno) %>% 
                                          filter(is.na(Comparison )))[, 1:4]
                                        

## parameter "compared_to_itself = true" means the two inputs were identical region lists 
# and will exclude comparison the same region to region itself in the process. 
DEG_to_DEG_distance_Both <- check_region_region_distance(Liver_DEG_list_Both_pos, 
                                                      Liver_DEG_list_Both_pos, compared_to_itself = TRUE)
DEG_to_DEG_distance_Pheno <- check_region_region_distance(Liver_DEG_list_Pheno_pos, 
                                                         Liver_DEG_list_Pheno_pos, compared_to_itself = TRUE)
nonDEG_to_DEG_distance_Both <- check_region_region_distance(Liver_nonDEG_list_Both_pos,
                                                      Liver_DEG_list_Both_pos, compared_to_itself = TRUE)
nonDEG_to_DEG_distance_Pheno <- check_region_region_distance(Liver_nonDEG_list_Pheno_pos,
                                                         Liver_DEG_list_Pheno_pos, compared_to_itself = TRUE)

saveRDS(DEG_to_DEG_distance_Both, file = paste0(dir_github_intermediate_files_fig3, 
                                                "DEG_to_DEG_distance_Both.rds"))
saveRDS(DEG_to_DEG_distance_Pheno, file = paste0(dir_github_intermediate_files_fig3,
                                                 "DEG_to_DEG_distance_Pheno.rds"))
saveRDS(nonDEG_to_DEG_distance_Both, file = paste0(dir_github_intermediate_files_fig3,
                                                   "nonDEG_to_DEG_distance_Both.rds"))
saveRDS(nonDEG_to_DEG_distance_Pheno, file = paste0(dir_github_intermediate_files_fig3,
                                                        "nonDEG_to_DEG_distance_Pheno.rds"))


```


### slncRNA and slncRNA
#### Intermediate file 3.9) slncRNA_to_slncRNA_distance_Both.rds
#### Intermediate file 3.10) slncRNA_to_slncRNA_distance_Pheno.rds
#### Intermediate file 3.11) nonslncRNA_to_slncRNA_distance_Both_reformated.rds
#### Intermediate file 3.12) nonslncRNA_to_slncRNA_distance_Pheno_reformated.rds

```{r}
# dir_github_script <- "G:/My Drive/github_script/" ## replace with your local directory
dir_lncRNA_count <- paste0(dir_github_script, "intermediate_files_prep/Figure3/")
dir_github_intermediate_files_fig3 <- paste0(dir_github_script, 
                                             "intermediate_files_prep/Figure3/collection_intermediate_files/")


Liver_slncRNA_XXfvsXYm <- readRDS(file = paste0(dir_lncRNA_count, "Liver_slncRNA_XXfvsXYm.rds"))
Liver_slncRNA_XYfvsXYm <- readRDS(file = paste0(dir_lncRNA_count, "Liver_slncRNA_XYfvsXYm.rds"))

##### check distance between nearest slncRNA
slncRNA_to_slncRNA_distance_Both <- check_region_region_distance(Liver_slncRNA_XXfvsXYm, 
                                                             Liver_slncRNA_XXfvsXYm, compared_to_itself = TRUE)
slncRNA_to_slncRNA_distance_Pheno <- check_region_region_distance(Liver_slncRNA_XYfvsXYm, 
                                                              Liver_slncRNA_XYfvsXYm, compared_to_itself = TRUE)

#### Heavy for a local computer to calculate. Prepared ahead of time on cluster.
# nonslncRNA_to_slncRNA_dist_Both <- check_region_region_distance(Liver_slncRNA_XXfvsXYm,
#                                                       Liver_nonslncRNA_XXfvsXYm, compared_to_itself = TRUE)
# nonslncRNA_to_slncRNA_dist_Pheno <- check_region_region_distance(Liver_slncRNA_XYfvsXYm,
#                                                       Liver_nonslncRNA_XYfvsXYm, compared_to_itself = TRUE)
nonslncRNA_to_slncRNA_dist_Both <- readRDS(file = paste0(dir_lncRNA_count, "nonslncRNA_to_slncRNA_dist_Both.rds"))
nonslncRNA_to_slncRNA_dist_Pheno <- readRDS(file = paste0(dir_lncRNA_count, "nonslncRNA_to_slncRNA_dist_Pheno.rds"))

saveRDS(slncRNA_to_slncRNA_distance_Both, file = paste0(dir_github_intermediate_files_fig3, 
                                                "slncRNA_to_slncRNA_distance_Both.rds"))
saveRDS(slncRNA_to_slncRNA_distance_Pheno, file = paste0(dir_github_intermediate_files_fig3,
                                                 "slncRNA_to_slncRNA_distance_Pheno.rds"))
saveRDS(nonslncRNA_to_slncRNA_dist_Both, file = paste0(dir_github_intermediate_files_fig3,
                                                   "nonslncRNA_to_slncRNA_dist_Both.rds"))
saveRDS(nonslncRNA_to_slncRNA_dist_Pheno, file = paste0(dir_github_intermediate_files_fig3,
                                                        "nonslncRNA_to_slncRNA_dist_Pheno.rds"))



```


### sDETE and sDEG
#### Intermediate file 3.13) all_DETE_DEG_distance_Both.rds
#### Intermediate file 3.14) all_DETE_DEG_distance_Pheno.rds
#### Intermediate file 3.15) random_draft_DETE_to_gene_distance_Both.rds
#### Intermediate file 3.16) random_draft_DETE_to_gene_distance_Pheno.rds

```{r}
# dir_github_script <- "G:/My Drive/github_script/" ## replace with your local directory
dir_to_Liver_TE_exp <- paste0(dir_github_script, "intermediate_files_prep/Figure3/")
dir_github_intermediate_files_fig3 <- paste0(dir_github_script, 
                                             "intermediate_files_prep/Figure3/collection_intermediate_files/")

## Distance from all genes to TE
#### All genes to all DETE
liver_diff_TE_annot_genic <- readRDS(file = paste0(dir_to_Liver_TE_exp, "liver_diff_TE_annot_genic.rds")) %>% filter(abs(log2FoldChange) >= 1)
liver_diff_TE_annot_genic_Both <- liver_diff_TE_annot_genic %>% filter(Comparison == "XX.F vs XY.M")
liver_diff_TE_annot_genic_Pheno <- liver_diff_TE_annot_genic %>% filter(Comparison == "XY.F vs XY.M")


#### Heavy for a local computer to calculate. Prepared ahead of time on cluster.
# DETE_all_gene_distance_Both <- check_region_region_distance(liver_diff_TE_annot_genic_Both, pos_of_gene_dedup)
# DETE_all_gene_distance_Pheno <- check_region_region_distance(liver_diff_TE_annot_genic_Pheno, pos_of_gene_dedup)
DETE_all_gene_distance_Both <- readRDS(file = paste0(dir_to_Liver_TE_exp,
              "DETE_all_gene_distance_Both.rds"))
DETE_all_gene_distance_Pheno <- readRDS(file = paste0(dir_to_Liver_TE_exp,
              "DETE_all_gene_distance_Pheno.rds"))



random_draft_DETE_to_gene_distance_Both <- sample(DETE_all_gene_distance_Both[[2]]$dist_check, 1000)
random_draft_DETE_to_gene_distance_Pheno <- sample(DETE_all_gene_distance_Pheno[[2]]$dist_check, 1000)

for(i in 1:10000){
   random_draft_DETE_to_gene_distance_Both <- (random_draft_DETE_to_gene_distance_Both + 
                                                  sample(DETE_all_gene_distance_Both[[2]]$dist_check, 1000))/2
   random_draft_DETE_to_gene_distance_Pheno <- (random_draft_DETE_to_gene_distance_Pheno + 
                                                  sample(DETE_all_gene_distance_Pheno[[2]]$dist_check, 1000))/2
}

sDETE_to_gene_distance_Both_reformated <- tibble(dist_check = random_draft_DETE_to_gene_distance_Both)  %>% 
                                              mutate(Group = "sDETE_to_random_gene(XX.FvsXY.M)") %>%
                                dplyr::select(c( "dist_check", "Group")) %>% 
                                    arrange(dist_check) %>% filter(dist_check > 0)

sDETE_to_gene_distance_Pheno_reformated <- tibble(dist_check = random_draft_DETE_to_gene_distance_Pheno)  %>% 
                                              mutate(Group = "sDETE_to_random_gene(XX.FvsXY.M)") %>%
  dplyr::select(c("dist_check", "Group")) %>% 
  arrange(dist_check) %>% filter(dist_check > 0)



## Distance from DEG to nearest DETE
Liver_DEG_list <- read_excel(paste0(dir_to_Liver_TE_exp, "Liver_DEG_recompiled.xlsx")) %>% mutate(gene = gene_symbol,
                                                                                         DEG = "Yes") %>% 
                                                                                  filter(abs(log_FC) > 1, log_CPM > 1, 
                                                                                         deseq.adj.pvalue < 0.05)
Liver_DEG_Both <- Liver_DEG_list %>% filter(Comparison == "XX.F vs XY.M")
Liver_DEG_Pheno <- Liver_DEG_list %>% filter(Comparison == "XY.F vs XY.M")


Liver_DEG_Both_pos <- left_join(Liver_DEG_Both, pos_of_gene_dedup %>% mutate()) %>% filter(!is.na(strand))
Liver_DEG_Pheno_pos <- left_join(Liver_DEG_Pheno, pos_of_gene_dedup %>% mutate()) %>% filter(!is.na(strand))


### DETE to DEG
liver_diff_TE_annot_genic <- readRDS(file = paste0(dir_to_Liver_TE_exp, "liver_diff_TE_annot_genic.rds"))
liver_diff_TE_annot_genic_Both <- liver_diff_TE_annot_genic %>% filter(Comparison == "XX.F vs XY.M") %>% 
                                          filter(abs(log2FoldChange) >= 1)
liver_diff_TE_annot_genic_Pheno <- liver_diff_TE_annot_genic %>% filter(Comparison == "XY.F vs XY.M")%>% 
                                          filter(abs(log2FoldChange) >= 1)

#### all DETE to DEG
mm10_rmsk_TE_strand <- read.table(file = paste0(dir_to_Liver_TE_exp, "mm10_rmsk_TE_strand_label.gtf"), 
                                  header = TRUE) %>% as_tibble() %>% mutate(chr = paste0("chr", chr))

liver_diff_TE_annot_genic_Both_strand <- left_join(liver_diff_TE_annot_genic_Both, mm10_rmsk_TE_strand)
all_DETE_DEG_distance_Pheno_strand <- left_join(liver_diff_TE_annot_genic_Pheno, mm10_rmsk_TE_strand)


all_DETE_DEG_distance_Both <- check_region_region_distance(liver_diff_TE_annot_genic_Both_strand,
                                                           Liver_DEG_Both_pos)
all_DETE_DEG_distance_Pheno <- check_region_region_distance(all_DETE_DEG_distance_Pheno_strand,  
                                                            Liver_DEG_Pheno_pos)


saveRDS(all_DETE_DEG_distance_Both, file = paste0(dir_github_intermediate_files_fig3, 
                                                "all_DETE_DEG_distance_Both.rds"))
saveRDS(all_DETE_DEG_distance_Pheno, file = paste0(dir_github_intermediate_files_fig3,
                                                 "all_DETE_DEG_distance_Pheno.rds"))
saveRDS(random_draft_DETE_to_gene_distance_Both, file = paste0(dir_github_intermediate_files_fig3,
                                                   "random_draft_DETE_to_gene_distance_Both.rds"))
saveRDS(random_draft_DETE_to_gene_distance_Pheno, file = paste0(dir_github_intermediate_files_fig3,
                                                        "random_draft_DETE_to_gene_distance_Pheno.rds"))



```


### sDETE and slncRnA
#### Intermediate file 3.17) slncRNA_to_sDETE_distance_Both.rds
#### Intermediate file 3.18) slncRNA_to_sDETE_distance_Pheno.rds
#### Intermediate file 3.19) random_draft_DETE_to_lncRNA_distance_Both.rds
#### Intermediate file 3.20) random_draft_DETE_to_lncRNA_distance_Pheno.rds

```{r}
# dir_github_script <- "G:/My Drive/github_script/" ## replace with your local directory
dir_to_Liver_TE_exp <- paste0(dir_github_script, "intermediate_files_prep/Figure3/")
dir_lncRNA_count <- paste0(dir_github_script, "intermediate_files_prep/Figure3/")
dir_github_intermediate_files_fig3 <- paste0(dir_github_script, 
                                             "intermediate_files_prep/Figure3/collection_intermediate_files/")


#### All lncRNA to all DETE
liver_sDETE_annot <- readRDS(file = paste0(dir_to_Liver_TE_exp, "liver_diff_TE_annot_genic.rds")) %>% filter(abs(log2FoldChange) >= 1)
liver_diff_TE_annot_genic_Both <- liver_sDETE_annot %>% filter(Comparison == "XX.F vs XY.M")
liver_diff_TE_annot_genic_Pheno <- liver_sDETE_annot %>% filter(Comparison == "XY.F vs XY.M")

#### Heavy for a local computer to calculate. Prepared ahead of time on cluster.
##### check distance between nearest slncRNA
# lncRNA_to_sDETE_distance_Both <- check_region_region_distance(liver_diff_TE_annot_genic_Both, 
#                                              Liver_all_lncRNA_pos, compared_to_itself = TRUE)
# lncRNA_to_sDETE_distance_Pheno <- check_region_region_distance(liver_diff_TE_annot_genic_Pheno, 
#                                               Liver_all_lncRNA_pos, compared_to_itself = TRUE)
lncRNA_to_sDETE_distance_Both <- readRDS(file = paste0(dir_lncRNA_count,
                                                     "lncRNA_to_sDETE_distance_Both.rds"))
lncRNA_to_sDETE_distance_Pheno <- readRDS(file = paste0(dir_lncRNA_count,
                                                        "lncRNA_to_sDETE_distance_Pheno.rds"))

#### Heavy for a local computer to calculate. Prepared ahead of time on cluster.
# slncRNA_to_sDETE_distance_Both <- check_region_region_distance(liver_diff_TE_annot_genic_Both, 
#                                              Liver_slncRNA_XXfvsXYm, compared_to_itself = TRUE)
# slncRNA_to_sDETE_distance_Pheno <- check_region_region_distance(liver_diff_TE_annot_genic_Pheno, 
#                                               Liver_slncRNA_XYfvsXYm, compared_to_itself = TRUE)
slncRNA_to_sDETE_distance_Both <- readRDS(file = paste0(dir_lncRNA_count,
                                                     "slncRNA_to_sDETE_distance_Both.rds"))
slncRNA_to_sDETE_distance_Pheno <- readRDS(file = paste0(dir_lncRNA_count,
                                                        "slncRNA_to_sDETE_distance_Pheno.rds"))



random_draft_DETE_to_lncRNA_distance_Both <- sample(lncRNA_to_sDETE_distance_Both[[2]]$dist_check, 1000)
random_draft_DETE_to_lncRNA_distance_Pheno <- sample(lncRNA_to_sDETE_distance_Pheno[[2]]$dist_check, 1000)

for(i in 1:10000){
   random_draft_DETE_to_lncRNA_distance_Both <- (random_draft_DETE_to_lncRNA_distance_Both + 
                                                  sample(lncRNA_to_sDETE_distance_Both[[2]]$dist_check, 1000))/2
   random_draft_DETE_to_lncRNA_distance_Pheno <- (random_draft_DETE_to_lncRNA_distance_Pheno + 
                                                  sample(lncRNA_to_sDETE_distance_Pheno[[2]]$dist_check, 1000))/2
}

sDETE_to_lncRNA_distance_Both_reformated <- tibble(dist_check =random_draft_DETE_to_gene_distance_Both)  %>% 
                                              mutate(Group = "sDETE_and_randomlncRNA(XX.FvsXY.M)") %>%
                                dplyr::select(c( "dist_check", "Group")) %>% 
                                    arrange(dist_check) %>% filter(dist_check > 0)

sDETE_to_lncRNA_distance_Pheno_reformated <- tibble(dist_check = random_draft_DETE_to_lncRNA_distance_Pheno)  %>% 
                                              mutate(Group = "sDETE_and_randomlncRNA(XX.FvsXY.M)") %>%
  dplyr::select(c("dist_check", "Group")) %>% 
  arrange(dist_check) %>% filter(dist_check > 0)

saveRDS(slncRNA_to_sDETE_distance_Both, file = paste0(dir_github_intermediate_files_fig3, 
                                                "slncRNA_to_sDETE_distance_Both.rds"))
saveRDS(slncRNA_to_sDETE_distance_Pheno, file = paste0(dir_github_intermediate_files_fig3,
                                                 "slncRNA_to_sDETE_distance_Pheno.rds"))
saveRDS(random_draft_DETE_to_lncRNA_distance_Both, file = paste0(dir_github_intermediate_files_fig3,
                                                   "random_draft_DETE_to_lncRNA_distance_Both.rds"))
saveRDS(random_draft_DETE_to_lncRNA_distance_Pheno, file = paste0(dir_github_intermediate_files_fig3,
                                                        "random_draft_DETE_to_lncRNA_distance_Pheno.rds"))


```





