## This script is to explain how intermediate files were prepared.
## Some intermediate files had a high demand for computing resources and were 
### separately prepared on cluster instead.


Figure 2. 

List of intermediate files
2.1) Brain_gene_data_filtered.rds
2.2) Heart_gene_data_filtered.rds
2.3) Lung_gene_data_filtered.rds
2.4) Liver_gene_data_filtered.rds
2.5) list_sDEG_four_organs_for_upsetPlot_XXfvsXYm.rds




### Figure 2.

#### Libraries and global variables
```{r}
## Libraries
suppressMessages(library(tidyverse))
suppressMessages(library( "DESeq2" ))
suppressMessages(library(ggplot2))
suppressMessages(library(ggfortify))
suppressMessages(library(readxl))
suppressMessages(library(dplyr))

## global variable
dir_github_script <- "G:/My Drive/github_script/" ## replace with your local directory


```


#### Intermediate file 1.1) Brain_gene_data_filtered.rds
#### Intermediate file 1.1) Heart_gene_data_filtered.rds
#### Intermediate file 1.1) Lung_gene_data_filtered.rds
#### Intermediate file 1.1) Liver_gene_data_filtered.rds

```{r}
# dir_github_script <- "G:/My Drive/github_script/"
dir_TE_exp <- paste0(dir_github_script, "intermediate_files_prep/Figure2/")
dir_gene_exp <- paste0(dir_github_script, "intermediate_files_prep/Figure2/")

gene_chr_pos <- read.csv(file = paste0(dir_TE_exp, "mm10_gene_loc_uniq.csv"), header = TRUE)
dir_github_intermediate_files_fig2 <- paste0(dir_github_script, 
                                             "intermediate_files_prep/Figure2/collection_intermediate_files/")

#### Four organ gene exp PCA
four_organ_gene_exp <- read.csv(paste0(dir_gene_exp, "Four_organs_gene_tb_raw.csv"))
#### To include HXX14 that was missing. 
heart_HXX14f <- read_excel(paste0(dir_gene_exp, "new_HXX14_counts.xlsx"))

#### Remove genes with extremely low coverage: read sums < 50 
four_organ_gene_exp_filtered <- four_organ_gene_exp[rowSums(four_organ_gene_exp[, 3:73 ])> 50,]

four_organ_gene_exp_filtered <- left_join(four_organ_gene_exp_filtered, gene_chr_pos)  %>% na.omit()

Sample.ID <- four_organ_gene_exp_filtered[,3:73] %>% names()
Organ <- c(rep("Liver", 21), rep("Brain", 18), rep("Heart", 16), rep("Lung", 16))
Group <- c(rep("XX.F", 4), rep("XY.M", 4), rep("XY.F", 4), rep("XO.F", 4), rep("XXpaf.F", 5),
           rep("XX.F", 4), rep("XY.M", 4), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 3),
           rep("XX.F", 2), rep("XY.M", 3), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 4),
           rep("XX.F", 3), rep("XY.M", 3), rep("XY.F", 4), rep("XO.F", 3), rep("XXpaf.F", 3))
gene.TE <- four_organ_gene_exp_filtered$Gene

four_organs_colors <- c("#FF7F0E", "#2CA02C", "#1F77B4", "#e377c2")

#### raw counts
four_organ_normalized_raw <- four_organ_gene_exp_filtered[, 3:73]
#### update to be norm counts, normalized by library depth of each sample.
four_organ_sum_depths <- four_organ_normalized_raw %>% lapply(., sum) %>% unlist()
four_organ_sum_depths_average <- four_organ_sum_depths  %>% mean()
for(i in 1:70){
  four_organ_normalized_raw[, i] <- (four_organ_normalized_raw[, i] * four_organ_sum_depths_average/four_organ_sum_depths[i]) %>% as.integer()
}

## To include chromosome information for gene
Gene_chr <- four_organ_gene_exp_filtered$chr
four_organ_normalized_raw_labeled <- cbind(four_organ_normalized_raw, Gene_chr)
#### X-linked
four_organ_normalized_raw_X <- four_organ_normalized_raw_labeled %>% filter(Gene_chr == "chrX") %>% dplyr::select(!"Gene_chr")
#### Autosomal
four_organ_normalized_raw_auto <- four_organ_normalized_raw_labeled %>% filter(Gene_chr != "chrX",
                                                                               Gene_chr != "chrY",
                                                                               Gene_chr != "chr1_GL456221_random",
                                                                               Gene_chr != "chr4_JH584293_random",
                                                                               Gene_chr != "chr4_JH584294_random",
                                                                               Gene_chr != "chrX_GL456233_random",
                                                              !is.na(Gene_chr)) %>% dplyr::select(!"Gene_chr")

### reformat the data for downstream analysis
four_organs_gene_reformatted_v2 <- cbind((t(four_organ_normalized_raw) ) %>% as_tibble(), 
                                         Sample.ID,
                                         Group,
                                         Organ) %>% data.frame()
four_organs_gene_reformatted_v2_auto <- cbind((t(four_organ_normalized_raw_auto) ) %>% as_tibble(), 
                                              Sample.ID,
                                              Group,
                                              Organ) %>% data.frame()
four_organs_gene_reformatted_v2_X <- cbind((t(four_organ_normalized_raw_X) ) %>% as_tibble(), 
                                           Sample.ID,
                                           Group,
                                           Organ) %>% data.frame()


rownames(four_organs_gene_reformatted_v2) <- four_organs_gene_reformatted_v2$Sample.ID


#### Brain
Brain_gene_data <- four_organs_gene_reformatted_v2 %>% as_tibble() %>%
  filter(Organ == "Brain")
length_gene_data_Brain <- length(Brain_gene_data)

## remove rows with only 0 as entry value.
Brain_gene_data_filtered <- Brain_gene_data[, c(which(colSums(Brain_gene_data %>% dplyr::select(where(is.numeric)))> 0),
                                                (length_gene_data_Brain-2):length_gene_data_Brain)] %>% data.frame()

rownames(Brain_gene_data_filtered) <- Brain_gene_data_filtered$Sample.ID

saveRDS(Brain_gene_data_filtered, file = paste0(dir_github_intermediate_files_fig2, "Brain_gene_data_filtered.rds"))


#### Heart
Heart_gene_data <- four_organs_gene_reformatted_v2 %>% as_tibble() %>%
  filter(Organ == "Heart", Sample.ID != "HXO6")

length_gene_data_Heart <- length(Heart_gene_data)

## remove rows with only 0 as entry value.
Heart_gene_data_filtered <- Heart_gene_data[, c(which(colSums(Heart_gene_data %>% dplyr::select(where(is.numeric)))> 0),
                                                (length_gene_data_Heart-2):length_gene_data_Heart)] %>% data.frame()

rownames(Heart_gene_data_filtered) <- Heart_gene_data_filtered$Sample.ID

saveRDS(Heart_gene_data_filtered, file = paste0(dir_github_intermediate_files_fig2, "Heart_gene_data_filtered.rds"))

#### Lung
Lung_gene_data <- four_organs_gene_reformatted_v2 %>% as_tibble() %>%
  filter(Organ == "Lung")
length_gene_data_Lung <- length(Lung_gene_data)


## remove rows with only 0 as entry value.
Lung_gene_data_filtered <- Lung_gene_data[, c(which(colSums(Lung_gene_data %>% dplyr::select(where(is.numeric)))> 0),
                                              (length_gene_data_Lung-2):length_gene_data_Lung)] %>% data.frame()
rownames(Lung_gene_data_filtered) <- Lung_gene_data_filtered$Sample.ID

saveRDS(Lung_gene_data_filtered, file = paste0(dir_github_intermediate_files_fig2, "Lung_gene_data_filtered.rds"))

#### Liver
Liver_gene_data <- four_organs_gene_reformatted_v2 %>% as_tibble() %>%
  filter(Organ == "Liver")
length_gene_data_Liver <- length(Liver_gene_data)

## remove rows with only 0 as entry value.
Liver_gene_data_filtered <- Liver_gene_data[, c(which(colSums(Liver_gene_data %>% dplyr::select(where(is.numeric)))> 0),
                                                (length_gene_data_Liver-2):length_gene_data_Liver)] %>% data.frame()

rownames(Liver_gene_data_filtered) <- Liver_gene_data_filtered$Sample.ID

saveRDS(Liver_gene_data_filtered, file = paste0(dir_github_intermediate_files_fig2, "Liver_gene_data_filtered.rds"))

```


```{r}
# dir_github_script <- "G:/My Drive/github_script/"
dir_gene_exp <- paste0(dir_github_script, "intermediate_files_prep/Figure2/")
dir_github_intermediate_files_fig2 <- paste0(dir_github_script, 
                                             "intermediate_files_prep/Figure2/collection_intermediate_files/")

sDEG_four_organs <- read_excel(path = paste0(dir_gene_exp, 
                                             "sDEG_four_organs_August08_2023.xlsx")) %>% 
                          mutate(Comparison = factor(Comparison),
                                 deseq2.p.value = as.double(deseq2.p.value))

sDEG_brain_XXfvsXYm <- sDEG_four_organs %>% filter(Organ == "Brain", Comparison == "XX.F vs XY.M")
sDEG_heart_XXfvsXYm <- sDEG_four_organs %>% filter(Organ == "Heart", Comparison == "XX.F vs XY.M")
sDEG_lung_XXfvsXYm <- sDEG_four_organs %>% filter(Organ == "Lung", Comparison == "XX.F vs XY.M")
sDEG_liver_XXfvsXYm <- sDEG_four_organs %>% filter(Organ == "Liver", Comparison == "XX.F vs XY.M")


list_sDEG_four_organs_for_upsetPlot_XXfvsXYm <- list("Liver XX.FvsXY.M" = sDEG_liver_XXfvsXYm$gene_symbol,
                                            "Heart XX.FvsXY.M" = sDEG_heart_XXfvsXYm$gene_symbol,
                                            "Brain XX.FvsXY.M" = sDEG_brain_XXfvsXYm$gene_symbol,
                                            "Lung XX.FvsXY.M" = sDEG_lung_XXfvsXYm$gene_symbol)


saveRDS(list_sDEG_four_organs_for_upsetPlot_XXfvsXYm, file = paste0(dir_github_intermediate_files_fig2,
                                       "list_sDEG_four_organs_for_upsetPlot_XXfvsXYm.rds"))


```

